% \documentclass{book}

\documentclass[12pt]{article}
\usepackage[pdfborder={0 0 0.5 [3 2]}]{hyperref}%
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}%
\usepackage[shortalphabetic]{amsrefs}%
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{amssymb}                
\usepackage{amsmath}                
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{bbm}
\usepackage[table,xcdraw]{xcolor}
\usepackage{tikz}
\usepackage{float}
\usepackage{booktabs}
\usepackage{svg}
\usepackage{mathtools}
\usepackage{cool}
\usepackage{url}
\usepackage{graphicx,epsfig}
\usepackage{makecell}
\usepackage{array}

\graphicspath{ {images/} }

\def\noi{\noindent}
\def\T{{\mathbb T}}
\def\R{{\mathbb R}}
\def\N{{\mathbb N}}
\def\C{{\mathbb C}}
\def\Z{{\mathbb Z}}
\def\P{{\mathbb P}}
\def\E{{\mathbb E}}
\def\Q{\mathbb{Q}}
\def\ind{{\mathbb I}}

\begin{document}

\title{}
\author{\vspace{-10ex} }

\begin{center}
{\LARGE APMA 1650 -- Homework 5}\\
\vspace{5mm}
{\large Due Thursday, July 21, 2016}\\
\vspace{5mm}
Homework is due during class or by 3:45 pm in the homework drop box in 182 George St.\\
Show all of your work used in deriving your solutions.
\end{center}

\begin{enumerate}

\item Suppose that the random variables $Y_1$ and $Y_2$ have joint density function:\
\[
f(y_1, y_2) = \begin{cases}
y_1 + y_2 & 0 \leq y_1 \leq 1, 0 \leq y_2 \leq 1 \\
0 & \text{otherwise}
\end{cases}
\]
\begin{enumerate}
\item Show this is a valid joint probability density. 
\item Find the marginal densities for $Y_1$ and $Y_2$.
\item Find the probability that $Y_1 < 1/2$ and $Y_2 > 1/2$.
\item Find the conditional density for $Y_1$ given $Y_2 = y_2$.
\item Find the probability that $Y_1 < 1/2$ given $Y_2 = 1/2$.
\end{enumerate}

\item Suppose that the random variables $Y_1$ and $Y_2$ have joint density function:\
\[
f(y_1, y_2) = \begin{cases}
c(1 - y_2) & 0 \leq y_1 \leq y_2 \leq 1 \\
0 & \text{otherwise}
\end{cases}
\]
\begin{enumerate}
\item Find the value of $c$ that makes this a valid joint probability density. 
\item Find the marginal densities for $Y_1$ and $Y_2$.
\item Find the conditional density for $Y_1$ given $Y_2 = y_2$.
\item Find the expected values $\E(Y_1)$ and $\E(Y_2)$.
\item Find the conditional expected value $\E(Y_1 | Y_2 = y_2)$.
\end{enumerate}

\item Suppose that the random variables $Y_1$ and $Y_2$ have joint density function:
\[
f(y_1, y_2) = \begin{cases}
e^{-(y_1 + y_2)} & y_1 > 0, y_2 > 0 \\
0 & \text{otherwise}
\end{cases}
\]
\begin{enumerate}
\item Find the marginal densities for $Y_1$ and $Y_2$. What kind of random variables are $Y_1$ and $Y_2$?
\item Find the conditional density of $Y_1$ given that $Y_2 = y_2$ for $y_2 > 0$.
\item Are $Y_1$ and $Y_2$ independent? Justify your answer.
\end{enumerate}

\item You are again the quality control manager for the Acme Widget Company. You have just launched a new line of MiniWidgets. Your MiniWidgets are produced by a MiniWidget machine. The MiniWidgets produced by the machine have masses which are normally distributed with a standard deviation of 0.2 grams. The machine can be adjusted so that the MiniWidgets it produces have an average mass of $\mu$ grams. What setting for $\mu$ should you use so that the masses of the MiniWidgets will exceed 10 grams at most 2\% of the time?

\item Let $X_1$ and $X_2$ be two independent geometric random variables, both with parameter $p$. Find a nice, closed-form formula for
\[
\P(X_1 = i | X_1 + X_2 = n)
\]
Your formula should not involve a sum. Hint: use the definition of conditional probability.
\end{enumerate}
\end{document}

