% \documentclass{book}

\documentclass[12pt]{article}
\usepackage[pdfborder={0 0 0.5 [3 2]}]{hyperref}%
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}%
\usepackage[shortalphabetic]{amsrefs}%
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{amssymb}                
\usepackage{amsmath}                
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{bbm}
\usepackage[table,xcdraw]{xcolor}
\usepackage{tikz}
\usepackage{float}
\usepackage{booktabs}
\usepackage{svg}
\usepackage{mathtools}
\usepackage{cool}
\usepackage{url}
\usepackage{graphicx,epsfig}
\usepackage{makecell}
\usepackage{array}

\def\noi{\noindent}
\def\T{{\mathbb T}}
\def\R{{\mathbb R}}
\def\N{{\mathbb N}}
\def\C{{\mathbb C}}
\def\Z{{\mathbb Z}}
\def\P{{\mathbb P}}
\def\E{{\mathbb E}}
\def\Q{\mathbb{Q}}
\def\ind{{\mathbb I}}

\graphicspath{ {images/} }

\begin{document}

\title{}
\author{\vspace{-10ex} }

\begin{center}
{\LARGE APMA 1650 -- Problem Session 2}\\
\vspace{5mm}
{\large Wednesday, July 20, 2016}\\
% \vspace{5mm}
\end{center}

There are seven problems on this sheet. There is no particular order to them, so I recommend that you work on the ones you find the most interesting.

\begin{enumerate}

\item For a certain section of a pine forest, the number of diseased trees per acre, $Y$ , has a Poisson distribution with mean $\lambda = 10$. The diseased trees are sprayed with an insecticide at a cost of \$3 per tree, plus a fixed overhead cost for equipment rental of \$50. Letting $C$ denote the total spraying cost for a randomly selected acre, find the expected value and variance for $C$. Within what interval would you expect $C$ to lie with probability of at least $0.75$?

\item Lifetimes of automotive tires are given in miles; higher performance tires are rated to last more miles. A manufacturer of tires wants to advertise a mileage interval that includes 90\% of the mileage on tires they manufacture. All they know is that, for a large number of tires tested, the mean mileage was 25,000 miles, and the standard deviation was 4000 miles. What interval would you suggest?

\item A machine used to fill cereal boxes dispenses, on average, $\mu$ ounces per box. Let $Y$ be the amount of ounces of cereal dispensed by the machine. The manufacturer wants $Y$ to be within 1 ounce of $\mu$ at least 75\% of the time. What is the largest value of the standard deviation $\sigma$ of $Y$ that can be tolerated if the manufacturer's objectives are met.

\item Beginning at 12:00 midnight, a call center is up for one hours and then down for two hours on a regular cycle. A person who is unaware of this schedule dials the center at a random time between 12:00 midnight at 5:00 am. What is the probability that the call center is up when the person's call comes in?

\item Weekly CPU time used by a hedge fund can be modeled by a random variable $Y$ which has probability density function (measured in hours) given by:
\[
f(y) = \begin{cases}
c y^2(4-y) & 0 \leq y \leq 4 \\
0 & \text{otherwise}
\end{cases}
\]
\begin{enumerate}
\item Find the value of $c$ such that this is a valid density function.
\item Find the expected value and variance of $Y$.
\item The CPU time costs the firm \$200 per hour. Find the expected value and variance of the weekly cost for CPU time.
\end{enumerate}

\item The magnitude of earthquakes can be modeled as an exponential distribution with mean 2.4 (as measured on the Richter scale). 
\begin{enumerate}
\item Find the probability that an earthquake will exceed 3.0 on the Richter scale.
\item Find the probability that an earthquake will fall between 2.0 and 3.0 on the Richter scale.
\end{enumerate}

\item Scores on an examination are (roughly) normally distributed with mean 78 and variance 36.
\begin{enumerate}
\item What is the probability that a student scores higher than 72?
\item Suppose that students in the top 20\% will receive an A on the exam? (This is now how I grade my exams.) What is the minimum score needed to receive an A?
\end{enumerate}

\end{enumerate}
\end{document}

